How to select predictive models for decision-making or causal inference?

### 2023-12-17 how_to_select_causal_models.tex
iter

Indeed, concluding on the effect of a treatment is a difficult
causal-inference task, as it can be easily compromised by confounding:
spurious associations between treatment
allocation and baseline health, \eg only prescribing a drug to mild cases
\cite{hernan_causal_2020,vanderweele2019principles}.
Predictive modeling bridges to causal inference theory under the name of
\emph{outcome models} (or G-computation, 
G-formula \cite{robins_role_1986}, Q-model
\cite{snowden_implementation_2011}, conditional mean regression
\cite{wendling_comparing_2018}). 
% Controlled allocation of treatment, as in randomized controlled trials (RCTs),
% alleviate this concern. Yet most machine-learning models are trained on
% \emph{observational} data, close to real-word practice
% \cite{black1996we,hernan_methods_2021} but challenging for causal inference.
Medical statistics and epidemiology have mostly used other
causal-inference methods, modeling treatment assignment
with propensity scores \cite{rosenbaum_central_1983,austin_moving_2015,casucci2018estimating,grose_use_2020}. Outcome modeling brings
the benefit of going
beyond average effects, estimating individualized or conditional average
treatment effects (CATE), central to precision medicine.
%
For this purpose, such methods are also invaluable on randomized trials
\cite{su2018random,lamont2018identification,hoogland2021tutorial}.

@@

Given complex health data, which model is to be most trusted to yield 
valid causal estimates needed to motivate individual treatment decisions?
As no single learner performs best on all data sets, there is a
pressing need for clear guidelines to select outcome models for causal
inference.

\subsection{Objectives and structure of the paper}

>>>>

Indeed, concluding on the effect of a treatment is a difficult
causal-inference task, as it can be easily compromised by confounding:
spurious associations between treatment
allocation and baseline health, \eg~only prescribing a drug to mild cases
\cite{hernan_causal_2020,vanderweele2019principles}.
Predictive modeling bridges to causal inference theory under the name of
\emph{outcome models} (or G-computation, 
G-formula \cite{robins_role_1986}, Q-model
\cite{snowden_implementation_2011}, conditional mean regression
\cite{wendling_comparing_2018}). 
% Controlled allocation of treatment, as in randomized controlled trials (RCTs),
% alleviate this concern. Yet most machine-learning models are trained on
% \emph{observational} data, close to real-word practice
% \cite{black1996we,hernan_methods_2021} but challenging for causal inference.
Medical statistics and epidemiology have mostly used other
causal-inference methods, modeling treatment assignment
with propensity scores \cite{rosenbaum_central_1983,austin_moving_2015,casucci2018estimating,grose_use_2020}. Outcome modeling brings
the benefit of going
beyond average effects, estimating individualized or conditional average
treatment effects (CATE), central to precision medicine.
%
For this purpose, such methods are also invaluable on randomized trials
\cite{su2018random,lamont2018identification,hoogland2021tutorial}.

@@

Given complex health data, which predictive model is to be most trusted to yield 
valid causal estimates needed to motivate individual treatment decisions?
As no single learner performs best on all data sets, there is a
pressing need for clear guidelines to select outcome models for causal
inference.

\paragraph{Objectives and structure of the paper}
